---
title: 'metaNanoPype reproducible report'
author: ""
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
bibliography: bibliography.bib
output: 
  html_document:
    toc: true 
    toc_float: true
    theme: united
    code_folding: hide
params: 
  input: input 
  output: output
---

<br>

<br>

---

<br>

<br>

## Report

<br>

This is a reproducible report of the taxonomic and diversity analyses produced by [metaNanoPype](https://antonioggsousa.github.io/metaNanoPype/).

<br>

The figures and tables displayed through the report can be downloaded in *pdf* and *tab-separated* format by clicking on the bottom left **Download plot** and **Download table** button that appears after each plot and table, respectively. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)
```

```{r klippy, echo=FALSE}
klippy::klippy(position = c('top', 'right'))
```

<br>

<br>

---

<br>

<br>

## About this report

<br>

This report was built with the `rmarkdown` R package (v.`r packageVersion("rmarkdown")`) [@rmarkdown_1; @rmarkdown_2; @rmarkdown_3]. The R version used was `r paste0(version$major, ".", version$minor)` [@r]. The taxonomic and divesity analyses were done with the `phyloseq` R package (v.`r packageVersion("phyloseq")`) [@phyloseq]. See the full list of packages and versions used at the end of this report - **R packages used and respective versions**.

```{r seed}

## Set seed 
set.seed(seed = 1024) # to keep reproducibility

```

```{r import packages, message=FALSE, warning=FALSE}

## Import packages 
library("DT", quietly = TRUE) 
library("ggplot2", quietly = TRUE)
library("phyloseq", quietly = TRUE)
source("./metaNanoPype_helper_funs.R")

```

<br>

<br>

---

<br>

<br>

## Taxonomy & Diversity

<br>

### Taxonomic table

<br>

```{r import & parse tax abund tbl}

# Import data
metananopype[["plots"]] <- metananopype[["params"]] <- metananopype[["data"]] <- metananopype <- list()
metananopype$params[["tax_path"]] <- params$input
metananopype$params[["out_path"]] <- params$output
files2import <- list.files(path = metananopype$params$tax_path, 
                           pattern = "\\.report", full.names = TRUE)
metananopype$params[["samples"]] <- gsub(pattern = ".report", replacement = "", 
                                         x = basename(files2import))
metananopype$data[["kraken2"]] <- lapply(X = setNames(files2import, metananopype$params$samples), 
                                         function(x, y = "\t", z = FALSE) read.table(file = x, sep = y, stringsAsFactors = z), 
                                         y = "\t", z = FALSE)

# Parse table
metananopype$data[["kraken2_parsed"]] <- lapply(metananopype$data$kraken2, function(x) parse_kraken2_report(x))
metananopype$data[["kraken2_parsed"]] <- lapply(metananopype$data$kraken2_parsed, 
                                                function(x, cols2sel = c(1,3)) x[,cols2sel], cols2sel = c(1,3))
for ( x in names(metananopype$data$kraken2_parsed)) colnames(metananopype$data$kraken2_parsed[[x]])[2] <- x
metananopype$data[["tax_tbl"]] <- merge_tax_tbl(list_tbls = metananopype$data$kraken2_parsed, merge_by = "Taxonomy")

# save table
if ( ! dir.exists(metananopype$params$out_path) ) dir.create(metananopype$params$out_path, recursive = TRUE)
write.table(x = cbind("Seq_id" = rownames(metananopype$data$tax_tbl),
                      metananopype$data$tax_tbl),
            file = paste(metananopype$params$out_path, "taxonomic_abundance_tbl.tsv", sep = "/"), 
            sep = "\t", row.names = FALSE)

```

```{r echo=FALSE}
datatable(data = metananopype$data$tax_tbl, extensions = "Buttons", caption = "Taxonomic abundance table.", 
          options = list(dom = "Blfrtip", buttons = c('csv', 'excel')))     
```

<br>

```{r echo=FALSE}
## Download the latest table
downloadthis::download_file(
  path = paste(metananopype$params[["out_path"]], "taxonomic_abundance_tbl.tsv", sep = "/"),
  output_name = "taxonomic_abundance_tbl.tsv",
  button_label = "Download table",
  button_type = "danger",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = TRUE
)
```

<br>

<br>

---

<br>

<br>

### Taxonomic barplot - Phylum

<br>

```{r import phyloseq}

## Import into phyloseq 
# 'tax_table'
full_tax <- metananopype$data$tax_tbl$Taxonomy
names(full_tax) <- row.names(metananopype$data$tax_tbl)
tax_tbl <- tax_to_physeq(full_tax = full_tax, split_by = "; ")
# 'otu_table
otu_tbl <- phyloseq::otu_table(object = as.matrix(metananopype$data$tax_tbl[,-1]), 
                               taxa_are_rows = TRUE)
# 'sam_data'
metadata <- data.frame("Samples" = colnames(metananopype$data$tax_tbl[,-1]), 
                       "SampleType" = colnames(metananopype$data$tax_tbl[,-1]))
row.names(metadata) <- metadata$Samples
metadata <- sample_data(metadata)
# phyloseq obj
metananopype$data[["physeq"]] <- phyloseq::phyloseq(otu_tbl, tax_tbl, metadata)

```

```{r tax barplot - phylum}

# Plot taxonomy
metananopype[["plots"]][["tax"]][["phyla"]] <- metananopype[["plots"]][["tax"]] <- list()
metananopype[["plots"]][["tax"]][["phyla"]] <- phyloseq::plot_bar(physeq = metananopype$data[["physeq"]], fill = "Phylum")
ggsave(filename = paste(metananopype$params[["out_path"]], "phyla_tax_barplot.pdf", sep = "/"), 
       plot = metananopype[["plots"]][["tax"]][["phyla"]])
print(metananopype[["plots"]][["tax"]][["phyla"]])

```

<br>

```{r echo=FALSE}
## Download the latest plot
downloadthis::download_file(
  path = paste(metananopype$params[["out_path"]], "phyla_tax_barplot.pdf", sep = "/"),
  output_name = "phyla_tax_barplot.pdf",
  button_label = "Download barplot",
  button_type = "danger",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = TRUE
)
```

<br>

<br>

---

<br>

<br>

### Alpha-diversity

<br>

```{r alpha-div}

# Plot alpha diversity
metananopype[["plots"]][["div"]][["alpha"]] <- list()
metananopype[["plots"]][["div"]][["alpha"]] <- phyloseq::plot_richness(physeq = metananopype$data[["physeq"]])
ggsave(filename = paste(metananopype$params[["out_path"]], "alpha_div_dotplot.pdf", sep = "/"), 
       plot = metananopype[["plots"]][["div"]][["alpha"]])
print(metananopype[["plots"]][["div"]][["alpha"]])

```

<br>

```{r echo=FALSE}
## Download the latest plot
downloadthis::download_file(
  path = paste(metananopype$params[["out_path"]], "alpha_div_dotplot.pdf", sep = "/"),
  output_name = "alpha_div_dotplot.pdf",
  button_label = "Download dotplot",
  button_type = "danger",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = TRUE
)
```

<br>

<br>

---

<br>

<br>

### Beta-diversity

<br>

```{r beta-div}

# Plot beta diversity
metananopype[["plots"]][["div"]][["beta"]] <- list()
metananopype$data[["ord"]] <- phyloseq::ordinate(physeq = metananopype$data[["physeq"]], 
                                                 method = "PCoA", distance = "bray")
metananopype[["plots"]][["div"]][["beta"]] <- phyloseq::plot_ordination(physeq = metananopype$data[["physeq"]],
                                                                        ordination = metananopype$data[["ord"]], 
                                                                        color = "Samples")
ggsave(filename = paste(metananopype$params[["out_path"]], "beta_div_plot.pdf", sep = "/"), 
       plot = metananopype[["plots"]][["div"]][["beta"]])
print(metananopype[["plots"]][["div"]][["beta"]])

```

<br>

```{r echo=FALSE}
## Download the latest plot
downloadthis::download_file(
  path = paste(metananopype$params[["out_path"]], "beta_div_plot.pdf", sep = "/"),
  output_name = "beta_div_plot.pdf",
  button_label = "Download plot",
  button_type = "danger",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = TRUE
)
```

<br>

<br>

---

<br>

<br>

#### R packages used and respective versions

```{r References, message=FALSE, warning=FALSE, paged.print=FALSE}

## R packages and versions used in these analyses

sessionInfo()

```

<br>

<br>

---   

<br>

<br>

**metaNanoPype** project page: <a href='https://antonioggsousa.github.io/metaNanoPype' target='_blank'>https://antonioggsousa.github.io/metaNanoPype</a>

<br>

<br>

---

<br>

<br>

## References

<br>

\printbibliography
